\section{CONCLUSION}
\label{sec:conclusion}

In this work, we presented a comprehensive investigation into real-time federated multi-task learning using Small Language Models across diverse Natural Language Understanding benchmarks. Our systematic evaluation across five architectures and multiple paradigms (Centralized vs.\ FL, IID vs.\ Non-IID) provides several critical insights for the deployment of decentralized NLU systems. Our results demonstrate that federated learning effectively preserves data privacy while significantly optimizing resource distribution. Specifically, we observed a reduction in per-client GPU resource consumption by \textbf{74--93\%} relative to centralized training. While the transition to a federated multi-task environment introduces a synchronization overhead of \textbf{2.2--3.7$\times$} in total training time due to communication rounds via WebSockets, the framework maintains competitive accuracy for classification tasks (SST-2 and QQP). However, task heterogeneity remains a significant challenge. We identified substantial negative transfer effects in regression tasks (STS-B) under high data heterogeneity (Non-IID), particularly for the smallest model variants (TinyBERT and BERT-Mini). Heterogeneity-aware aggregation and Low-Rank Adaptation (LoRA) proved essential, with LoRA achieving over \textbf{90\%} reduction in communication volume while partially recovering performance in complex task regimes.
Among the investigated SLM architectures, \textbf{DistilBERT} consistently provided the most robust balance between linguistic performance and system efficiency, showing superior resilience to the negative transfer effects inherent in multi-task federated learning. For extreme edge scenarios where memory is the primary bottleneck, \textbf{MiniLM} emerged as a viable alternative, though with a stricter trade-off in regression stability.

Future research will focus on developing adaptive task-sharing strategies that dynamically adjust parameter weights to further mitigate negative transfer in multi-head architectures. Additionally, we plan to explore personalized federated learning techniques to better accommodate extreme label skew and the integration of differentially private mechanisms within the real-time WebSocket communication layer to fortify privacy guarantees without compromising the efficiency gains observed in this study.
