\documentclass[conference]{IEEEtran}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}

\begin{document}

\title{Federated Learning with Small Language Models: Multi-Task Learning and Data Heterogeneity Analysis}

\author{
    \IEEEauthorblockN{Author Name(s)}
    \IEEEauthorblockA{Institution\\
    Email: author@example.com}
}

\maketitle

\begin{abstract}
This paper presents a comprehensive analysis of federated learning with small language models (SLMs) in multi-task learning scenarios under data heterogeneity. We evaluate the performance of five different model architectures across three NLU benchmarks (SST-2, QQP, STS-B) under centralized and federated paradigms, comparing single-task versus multi-task learning approaches. Our experiments reveal significant insights into task transfer effects, privacy-performance trade-offs, and efficiency considerations for practical deployment. Results show that while federated learning introduces 2.4-3.9x training overhead, it maintains competitive accuracy for classification tasks and reduces individual client resource usage by 77-93\%. Multi-task learning exhibits negative transfer effects in 85\% of configurations, with performance degradation ranging from -1.93\% to -26.58\% depending on task similarity and model capacity.
\end{abstract}

\begin{IEEEkeywords}
Federated Learning, Small Language Models, Multi-Task Learning, Data Heterogeneity, Privacy-Preserving Machine Learning
\end{IEEEkeywords}

\input{experiments_section}

\section{Conclusion}

This study provides a comprehensive analysis of federated learning with small language models in multi-task scenarios. Our key contributions include:

\begin{itemize}
    \item Systematic evaluation of five SLM architectures across three NLU benchmarks
    \item Detailed analysis of single-task vs multi-task learning in federated settings
    \item Quantification of privacy-performance trade-offs and efficiency metrics
    \item Practical insights for real-world deployment of federated systems
\end{itemize}

\textbf{Main Findings:}
\begin{enumerate}
    \item Federated learning maintains competitive performance for classification tasks with significant resource efficiency gains
    \item Multi-task learning shows limited benefits in federated scenarios due to negative transfer effects
    \item Model architecture selection crucial for balancing performance and efficiency
    \item Data heterogeneity significantly impacts convergence and final performance
\end{enumerate}

\textbf{Future Work:}
\begin{itemize}
    \item Exploration of adaptive task sharing strategies
    \item Investigation of personalized federated learning approaches
    \item Development of efficient communication protocols for multi-task FL
    \item Extension to more diverse task combinations and model architectures
\end{itemize}

\begin{thebibliography}{9}
\bibitem{ref1}
McMahan, B., et al. (2017). Communication-Efficient Learning of Deep Networks from Decentralized Data. \emph{AISTATS}.

\bibitem{ref2}
Devlin, J., et al. (2019). BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding. \emph{NAACL}.

\bibitem{ref3}
Wang, A., et al. (2019). GLUE: A Multi-Task Benchmark and Analysis Platform for Natural Language Understanding. \emph{ICLR}.

\bibitem{ref4}
Caruana, R. (1997). Multitask Learning. \emph{Machine Learning}, 28(1), 41-75.

\bibitem{ref5}
Kairouz, P., et al. (2021). Advances and Open Problems in Federated Learning. \emph{Foundations and Trends in Machine Learning}.

\bibitem{ref6}
Hsu, C., et al. (2020). Measuring the Effects of Non-Identical Data Distribution on Federated Learning. \emph{arXiv preprint}.

\bibitem{ref7}
Sanh, V., et al. (2019). DistilBERT, a distilled version of BERT: smaller, faster, cheaper and lighter. \emph{arXiv preprint}.

\bibitem{ref8}
Yang, Y., et al. (2020). Federated Learning Based on Dynamic Regularization. \emph{ICLR}.

\bibitem{ref9}
Wang, X., et al. (2021). FedBN: Federated Learning on Non-IID Data with Local Batch Normalization. \emph{arXiv preprint}.
\end{thebibliography}

\end{document}
