\begin{thebibliography}{31}
\bibitem{devlin2018bert}
			Devlin, J., Chang, M.-W., Lee, K., Toutanova, K. (2019). 
			BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding. 

\bibitem{gdpr}
	General Data Protection Regulation (GDPR), Accessed: Aug. 31, 2025. [Online]. Available: \url{https://gdpr-info.eu}

\bibitem{mcmahan2017communication}
	H. McMahan, E. Moore, D. Ramage, S. Hampson, and B. Arcas, ``Communication-efficient learning of deep networks from decentralized data,'' in \emph{Proceedings of AISTATS}, 2017, pp. 1273--1282.
	
\bibitem{yang2019federated}
	Q. Yang, Y. Liu, T. Chen, and Y. Tong, ``Federated machine learning: Concept and applications,'' \emph{ACM Transactions on Intelligent Systems and Technology}, vol. 10, no. 2, pp. 1--19, 2019.

\bibitem{liu2019roberta}
			Liu, Y., Ott, M., Goyal, N., Du, J., Joshi, M., Chen, D., Levy, O., Lewis, M., Zettlemoyer, L., Stoyanov, V. (2019). 
			RoBERTa: A Robustly Optimized BERT Pretraining Approach. 
			\textit{arXiv preprint arXiv:1907.11692}. \url{https://arxiv.org/abs/1907.11692}
						
\bibitem{li2020challenges}
T.~Li, A.~K. Sahu, A.~Talwalkar, and V.~Smith.
Federated learning: Challenges, methods, and future directions.
\emph{IEEE Signal Processing Magazine}, 37(3):50--60, 2020.

\bibitem{mcmahan2017fedavg}
H.~B. McMahan, E.~Moore, D.~Ramage, S.~Hampson, and B.~A. y Arcas.
Communication-efficient learning of deep networks from decentralized data.
In \emph{AISTATS}, 2017.

\bibitem{smith2017mocha}
V.~Smith, C.-K. Chiang, M.~Sanjabi, and A.~Talwalkar.
Federated multi-task learning.
In \emph{NeurIPS}, 2017.

\bibitem{chen2024fedbone}
Y.~Chen, T.~Zhang, X.~Jiang, Q.~Chen, C.~Gao, and W.~Huang.
FedBone: Towards Large-Scale Federated Multi-Task Learning.
\emph{Journal of Computer Science and Technology}, 1040-1057, 2024.

\bibitem{lin2021fednlp}
B.~Y.~Lin, C.~He, Z.~Ze, H.~Wang, Y.~Hua, C.~Dupuy, R.~Gupta, M.~Soltanolkotabi, X.~Ren, and S.~Avestimehr.
FedNLP: Benchmarking Federated Learning Methods for Natural Language Processing Tasks.
\emph{Findings of the Association for Computational Linguistics: NAACL 2022}, 157-175, 2022.

\bibitem{li2020fedprox}
T.~Li, A.~K. Sahu, M.~Zaheer, M.~Sanjabi, A.~Talwalkar, and V.~Smith.
Federated optimization in heterogeneous networks.
In \emph{MLSys}, 2020.

\bibitem{zhou2025mfed}
J.~Zhou, W.~Bao, J.~Wang, D.~Zhang, X.~Zhang, and Y.~Zhang.
Multi-task federated learning with encoder–decoder structure: enabling collaborative learning across different tasks.
\emph{International Journal of Machine Learning and Cybernetics}, 10403-10420, 2025.

\bibitem{elbakary2025mira}
A.~Elbakary, C.~Ben Issaid, T.~ElBatt, K.~Seddik, and M.~Bennis.
MIRA: A Method of Federated Multi-Task Learning for Large Language Models.
\emph{IEEE Networking Letters}, 171-175, 2025.

\bibitem{sanh2019distilbert}
V.~Sanh, L.~Debut, J.~Chaumond, and T.~Wolf.
DistilBERT, a distilled version of BERT: smaller, faster, cheaper and lighter.
\emph{arXiv preprint arXiv:1910.01108}, 2019.

\bibitem{turc2019compact}
I.~Turc, M.-W. Chang, K.~Lee, and K.~Toutanova.
Well-read students learn better: On the importance of pre-training compact models.
\emph{arXiv preprint arXiv:1908.08962}, 2019.

\bibitem{wang2020minilm}
W.~Wang, F.~Wei, L.~Dong, H.~Bao, N.~Yang, and M Zhou.
MiniLM: Deep self-attention distillation for task-agnostic compression of pre-trained transformers.
In \emph{NeurIPS}, 2020.

\bibitem{jiao2019tinybert}
X.~Jiao, Y.~Yin, L.~Shang, X.~Jiang, X.~Chen, L.~Li, F.~Wang, and Q.~Liu.
TinyBERT: Distilling BERT for Natural Language Understanding.
In \emph{Findings of the Association for Computational Linguistics: EMNLP 2020}, 4163-4174, 2020.

\bibitem{karimireddy2020scaffold}
S.~P. Karimireddy, S.~Kale, M.~Mohri, et al.
SCAFFOLD: Stochastic controlled averaging for federated learning.
In \emph{ICML}, 2020.

\bibitem{yu2020pcgrad}
T.~Yu, S.~Kumar, A.~Gupta, S.~Levine, K.~Hausman, and C.~Finn.
Gradient surgery for multi-task learning.
\emph{arXiv preprint arXiv:2001.06782}, 2020.

\bibitem{houlsby2019parameter}
N.~Houlsby, A.~Giurgiu, S.~Jastrzebski, et al.
Parameter-efficient transfer learning for NLP.
In \emph{ICML}, 2019.

\bibitem{pfeiffer2020adapterhub}
J.~Pfeiffer, A.~Rücklé, C.~Poth, A.~Kamath, I.~Vulić, S.~Ruder, K.~Cho, and I.~Gurevych.
AdapterHub: A Framework for Adapting Transformers.
In \emph{Proceedings of the 2020 Conference on Empirical Methods in Natural Language Processing: System Demonstrations}, 46-54, 2020.

\bibitem{lan2019albert}
Z.~Lan, M.~Chen, S.~Goodman, K.~Gimpel, P.~Sharma, and R.~Soricut.
ALBERT: A Lite BERT for self-supervised learning of language representations.
In \emph{ICLR}, 2020.

\bibitem{wang2025edgedistilbert}
S.~Wang, et al.
EdgeDistilBERT: A lightweight multi-task federated learning framework for resource-constrained devices.
\emph{arXiv preprint}, 2025.

\bibitem{li2021fedbn}
X.~Li, M.~Jiang, X.~Zhang, et al.
FedBN: Federated learning on non-IID data with batch normalization.
In \emph{ICLR}, 2021.

\bibitem{huo2022fedpara}
Z.~Huo, M.~Bin~Ahmad, S.~Guu, et al.
FedPara: Low-rank Hadamard product for communication-efficient federated learning.
In \emph{ICLR}, 2022.

\bibitem{khan2024systematic}
Y.~Khan, D.~Sánchez, and J.~Domingo-Ferrer.
Federated learning-based natural language processing: a systematic literature review.
\emph{Artificial Intelligence Review}, 2024.

\bibitem{devlin2018bert}
J.~Devlin, M.-W. Chang, K.~Lee, and K.~Toutanova.
BERT: Pre-training of deep bidirectional transformers for language understanding.
In \emph{Proceedings of the 2019 conference of the North American chapter of the association for computational linguistics: human language technologies, volume 1 (long and short papers)}, pages 4171--4186, 2019.

\bibitem{reddi2021adaptive}
S.~Reddi, Z.~Charles, M.~Zaheer, et al.
Adaptive federated optimization.
In \emph{ICLR}, 2021.

\bibitem{hu2021lora}
E.~J. Hu, Y.~Shen, P.~Wallis, Z.~Allen-Zhu, Y.~Li, S.~Wang, L.~Wang, and W.~Chen.
LoRA: Low-rank adaptation of large language models.
In \emph{ICLR}, 2022.

\bibitem{wang2018glue}
A.~Wang, A.~Singh, J.~Michael, F.~Hill, O.~Levy, and S.~R. Bowman.
GLUE: A multi-task benchmark and analysis platform for natural language understanding.
In \emph{ICLR}, 2019.

\bibitem{socher2013sst}
R.~Socher, A.~Perelygin, J.~Wu, J.~Chuang, C.~D.~Manning, A.~Ng, and C.~Potts.
Recursive Deep Models for Semantic Compositionality Over a Sentiment Treebank.
In \emph{Proceedings of the 2013 Conference on Empirical Methods in Natural Language Processing}, 1631-1642, 2013.

\bibitem{vanrijsbergen1979}
C.~J. Van Rijsbergen.
\emph{Information Retrieval}.
Butterworth-Heinemann, 2nd edition, 1979.

\bibitem{sharma2019qqp}
L.~Sharma, L.~Graesser, N.~Nangia, and U.~Evci.
Natural language understanding with the quora question pairs dataset.
\emph{arXiv preprint arXiv:1907.01041}, 2019.

\bibitem{cer2017semeval}
D.~Cer, M.~Diab, E.~Agirre, I.~Lopez-Gazpio, and L.~Specia.
SemEval-2017 Task 1: Semantic Textual Similarity Multilingual and Crosslingual Focused Evaluation.
In \emph{Proceedings of the 11th International Workshop on Semantic Evaluation (SemEval-2017)}, 1-14, 2017.

\bibitem{pearson1895}
K.~Pearson.
VII. Note on regression and inheritance in the case of two parents.
\emph{Proceedings of the Royal Society of London}, 58:240--242, 1895.

\bibitem{spearman1904}
C.~Spearman.
The proof and measurement of association between two things.
\emph{The American Journal of Psychology}, 15(1):72--101, 1904.

\end{thebibliography}
