% \section{RELATED WORK}
% \label{sec:related_work}

% \subsection{Federated Learning for Natural Language Processing}
% FL has been applied to several NLP/NLU scenarios, including text classification and personalized language services where raw text cannot be centralized \cite{mcmahan2017fedavg,li2020fedprox}. Foundational optimization methods such as FedAvg established the standard client-update and server-aggregation loop, while FedProx introduced regularization to improve robustness under client drift \cite{mcmahan2017fedavg,li2020fedprox}.

% For heterogeneous settings, federated multi-task approaches such as MOCHA show that modeling inter-task relatedness can improve statistical efficiency and robustness to systems issues \cite{smith2017mocha}. More recent methods, including FedBone, extend this direction toward large-scale heterogeneous multi-task systems by explicitly separating shared and task-specific components \cite{chen2024fedbone}.

% Despite this progress, many FL-for-NLP studies still assume limited task diversity or focus primarily on single-task personalization.

% \subsection{Multi-Task Learning in NLP}
% MTL in NLP usually follows two paradigms: hard parameter sharing and soft parameter sharing \cite{smith2017mocha,yu2020pcgrad}. Hard sharing uses a common encoder and separate task heads, reducing overfitting and training cost. Soft sharing keeps task-specific parameters while encouraging similarity through regularization or constrained interaction.

% MTL improves representation learning by transferring useful linguistic features across tasks. However, it also introduces known difficulties, including task imbalance (dominant tasks overpower smaller ones) and negative transfer (one task degrades another). These issues become more severe under federated training because client participation is partial and data distributions vary across rounds.

% \subsection{Small Language Models for Efficient NLP}
% SLMs such as DistilBERT, MiniLM, and ALBERT demonstrate that careful compression and distillation can preserve strong performance while reducing parameter count and inference cost \cite{sanh2019distilbert,wang2020minilm,lan2019albert}. Common reduction strategies include knowledge distillation, parameter sharing, low-rank factorization, and architectural simplification.

% These properties make SLMs highly suitable for edge and distributed learning: they reduce transmission overhead, lower on-device training cost, and enable faster federated iteration cycles. In federated multi-task environments, SLMs also make it easier to scale to more clients and tasks under fixed system budgets.

% \subsection{Federated Learning under Data Heterogeneity}
% Non-IID data remains one of the core bottlenecks in FL. Client-specific language patterns, domain shifts, and label skew can bias local updates and destabilize global optimization. In heterogeneous NLU deployments, this challenge is compounded by cross-client task diversity, where clients may optimize different objectives with partially shared representations.

% Prior work reports that heterogeneity can hurt both convergence speed and final generalization if aggregation ignores task structure. Approaches such as personalized layers, proximal regularization, gradient conflict mitigation, and encoder--decoder decomposition have been explored to reduce this mismatch \cite{li2020fedprox,yu2020pcgrad,zhou2025mfed}.

% \subsection{Research Gap}
% Existing studies provide strong building blocks but still leave an important gap for practical NLU systems:
% \begin{itemize}
%     \item Lack of real-time FL frameworks: most methods are designed for periodic offline rounds rather than continuous adaptation.
%     \item Limited integration of FL + MTL + SLM: many works treat efficiency, heterogeneity, and multi-task transfer separately.
%     \item Insufficient heterogeneity analysis in NLU FL: current evaluations often miss how task diversity, data skew, and communication constraints interact.
% \end{itemize}

% This gap motivates a unified real-time federated multi-task approach that combines efficient language modeling with heterogeneity-aware optimization for distributed NLU.

\section{RELATED WORK}
\label{sec:related_work}

This section reviews the foundational research and current advancements in Federated Learning for NLP, MTL, and the development of SLMs We evaluate how these fields intersect to enable efficient distributed learning and identify the research gaps that our proposed real-time framework aims to address.

\subsection{Federated Learning for Natural Language Processing}
There are many examples of FL utilized in NLP/NLU applications such as text classification and personalized language services where raw text cannot be aggregated \cite{mcmahan2017fedavg,li2020fedprox}. Basic optimization methods like FedAvg had set up the typical client-update-server-aggregation loop, and FedProx added regularization to enable better behavior against client drift \cite{mcmahan2017fedavg,li2020fedprox}. Advanced methods like SCAFFOLD \cite{karimireddy2020scaffold} further address gradient heterogeneity to ensure stable convergence. Federated multi-task approaches, such as MOCHA, demonstrate that inter-task relatedness can be made more powerful under heterogeneous setups by improving statistical efficiency and robustness to system issues \cite{smith2017mocha}. Newer approaches such as FedBone take this further by explicitly separating shared from task-dependent elements in large heterogeneous multi-task architectures \cite{chen2024fedbone}. Notwithstanding this progress, a large number of FL-for-NLP studies rely on limited task diversity or mainly rely solely on single-task user personalization.

\subsection{Multi-task learning NLP}
MTL in NLP generally employs one of the two paradigms: hard parameter sharing and soft parameter sharing \cite{smith2017mocha,yu2020pcgrad}. Hard sharing employs a common encoder, instead of a separate task head for each instance of the model, saving the model from overfitting and training cost. Soft sharing retains task-specific factors but promotes similarity either by regularization or by limiting interaction. The adoption of parameter-efficient fine-tuning (PEFT), such as adapters \cite{houlsby2019parameter} and modular shared frameworks like AdapterHub \cite{pfeiffer2020adapterhub}, has further enabled efficient multi-task adaptation with minimal overhead. MTL improves representation learning by transfer of useful linguistic features between tasks. On the other hand, it also comes with known drawbacks, such as task imbalance (over-representation of dominant tasks with lower-level tasks) and negative transfer (one task harms others). These problems are more pronounced in the context of federated training because client involvement is partial, and the balance of data distribution between rounds is not consistent.

\subsection{NLP and Small Language Models}
SLMs like DistilBERT, MiniLM, and ALBERT have been shown that careful compression and distillation can preserve good performance with a cost between parameter count and inference \cite{sanh2019distilbert,wang2020minilm,lan2019albert}. The common reduction methods are knowledge distillation, parameter sharing, low-rank factorization, and architectural simplification. These capabilities make those SLMs very appropriate for edge and distributed learning: they decrease the transmission overhead, reduce on-device training cost, and achieve faster federated iteration cycles. Recent innovations like EdgeDistilBERT \cite{wang2025edgedistilbert} further push the boundaries of lightweight multi-task FL on edge devices. In settings with multistask federated solutions, SLMs also facilitate scaling onto larger numbers of clients and tasks on fixed resource budgets for the system.

\subsection{Federated Learning in Heterogeneous Data}
Non-IID data is still one of the major bottlenecks in FL. Client-dependent language patterns, domain shifts, and label bias may lead to biased updates at the local level and may disrupt global optimization at the global level. This problem is aggravated in heterogeneous NLU deployments where the difference between client tasks exists (cross-client task diversity) whereby the clients can perform partial representations for optimizing different objectives. Strategies such as FedBN \cite{li2021fedbn} and low-rank parameterizations like FedPara \cite{huo2022fedpara} have been investigated to mitigate these heterogeneity effects. As previous work report, aggregation without attention to the structure of a task will degrade velocity and final generalization, whereas when heterogeneity is mixed up (thereby aggregation). Since convergence speed can harm fine generalization and final speed depending on task structure, techniques such as personalized layers, proximal regularization, gradient conflict reducing, and encoder--decoder decomposition have been proposed to bridge this gap \cite{li2020fedprox,yu2020pcgrad,zhou2025mfed}.

\subsection{Research Gap}
Current studies offer powerful foundations, and still constitute, however, an important gap to be filled in terms of practical NLU systems, as identified in recent systematic reviews \cite{khan2024systematic}:
\begin{itemize}
    \item No real-time FL-based libraries: the majority of approaches only adopt offline versions for short time periods rather than for real-time adaptation.
    \item Inconsistent and partial inclusion of FL + MTL + SLM: the efficiency, heterogeneity and multi task-transfer of the models are frequently reported in different ways in papers.
    \item Insufficiently analysis heterogeneity in NLU FL: the present work tends to overlook how task diversity and the relationship between task diversity, data skew, and communication limitations work together.
\end{itemize}
This gap spurs a unified, real-time federated multi-task architecture comprising efficient language modeling and heterogeneous optimization for distributed NLU, coupled with optimization awareness.