% \section{PROBLEM DEFINITION}  
% \label{sec:problem_definition}

% This section formalizes the RT-FedMTL environment for NLU. We define the mathematical framework for distributed multi-task optimization, address the challenges of data and system heterogeneity, and establish the global objective for collaborative model training across private edge clients.

% \subsection{Formalization of RT-FedMTL for NLU}
% We consider a RT-FedMTL environment based on NLU in which training proceeds over communication rounds $t=1,2,\dots$ \cite{mcmahan2017fedavg,smith2017mocha,chen2024fedbone}. During each round, a subset of clients is selected, local updates are performed on private data streams, and parameters are pushed to a central server for aggregation. Unlike offline FL, client data distributions may vary over time; therefore, both local goals and optimum parameters are time-dependent. Let $\mathcal{C}=\{1,\dots,C\}$ be the set of clients and $\mathcal{K}=\{1,\dots,K\}$ be the set of NLU tasks. Each client $c\in\mathcal{C}$ is associated with one primary task $\tau(c)\in\mathcal{K}$ (or a small task subset), where new samples are continuously added from a local stream. The objective is to learn a federated model that maintains cross-task transfer quality while respecting task-level specialization under non-IID and time-varying data.

% \subsection{Data and Task Heterogeneity in Clients}  
% Heterogeneity comes in two forms:  
% \begin{itemize}  
% \item Heterogeneity of data: the local distributions in clients $c_i,c_j$ will be different (i.e. $P_{c_i}(x,y)\neq P_{c_j}(x,y)$ (label skew, domain shift, vocabulary differences).  
% \item Task heterogeneity: clients may optimize different losses and output spaces according to $\tau(c)$, yielding dissimilar gradients across tasks.  
% \end{itemize}  
% Moreover, real-time environments have client availability and update frequency asynchronous, leading to system heterogeneity and stale updates \cite{mcmahan2017fedavg,li2020fedprox,chen2024fedbone}. Thus, the optimization process must account for partial participation and non-uniform local progress.  

% \subsection{Optimization Objective for Federated Multi-Task Learning}  
% Let $\ell_{\tau(c)}(\theta_e,\theta_{\tau(c)};\xi)$ be the task-specific loss cost for sample $\xi\in\mathcal{D}_c^t$. The global objective at round $t$ is:  
% \begin{equation}  
% \begin{aligned}  
% \min_{\theta_e,\{\theta_k\}} \, F_t(\Theta)  
% &= \sum_{c=1}^{C} p_c^t \, \mathbb{E}_{\xi\sim\mathcal{D}_c^t}  
% \big[\ell_{\tau(c)}(\theta_e,\theta_{\tau(c)};\xi)\big] \\  
% &\quad + \lambda \, \Omega(\{\theta_k\}),  
% \end{aligned}  
% \end{equation}  
% where $p_c^t$ represents the weight of the aggregation (e.g., the weight per unit local sample), $\Omega$ is a regularizer to promote a stable cross-task learning and $\lambda\ge0$ controls the sharing-specialization trade-off. After localized updates, the server aggregates common encoder parameters by weighted averaging,  
% \begin{equation}  
% \theta_e^{t+1} = \sum_{c\in\mathcal{C}_t} \alpha_c^t \, \theta_{e,c}^{t+1}.  
% \end{equation}  
% and updates custom task-specific parameters, as a result of task-wise aggregation,  
% \begin{equation}  
% \theta_k^{t+1} = \sum_{c\in\mathcal{C}_t: \, \tau(c)=k} \beta_c^t \, \theta_{k,c}^{t+1}.  
% \end{equation}  
% In $\mathcal{C}_t$, the client subset participate at round $t$ and $\alpha_c^t,\beta_c^t$ are normalized aggregation coefficients \cite{mcmahan2017fedavg,smith2017mocha,zhou2025mfed}. With privacy and heterogeneity requirements in place, the framework captures RT-FedMTL in real time by jointly optimizing both shared representations and task-specific parameters.

\section{PROBLEM DEFINITION}  
\label{sec:problem_definition}  
This section introduces the official RT-FedMTL environment for NLU. We define the formal mathematical structure for distributed multi-task optimization work, tackle the issues of data and system heterogeneity, and set the global goal for training collaboration models across private edge clients.

\subsection{RT-FedMTL Formalization for NLU}  
We model a NLU-based RT-FedMTL system in which training is carried out across communication rounds $t=1,2,\dots$ \cite{mcmahan2017fedavg,smith2017mocha,chen2024fedbone}. In each round, a subset of clients is chosen, local updates are made to private data streams, and parameters are returned to a central server for aggregation. As opposed to offline FL, the client data distributions can vary over time. Hence, both local goals and optimum parameters are time-dependent. Let $\mathcal{C}=\{1,\dots,C\}$ represent the set of clients and $\mathcal{K}=\{1,\dots,K\}$ represent the set of NLU tasks. Each client $c\in\mathcal{C}$ is assigned to only one main task $\tau(c)\in\mathcal{K}$ (or a small subset of tasks), where new samples are added to the local stream perpetually. The goal is to learn a federated model with cross-task transfer quality but task-level specialization in the absence of IID and time-varying data.

\subsection{Distribution of Clients and the Data}  
Heterogeneity comes in two forms:  
\begin{itemize}  
\item Heterogeneity of Data: the local distributions will differ across clients $c_i,c_j$ (i.e. $P_{c_i}(x,y)\neq P_{c_j}(x,y)$) (label skew, domain shift, vocabulary differences).  
\item Task heterogeneity: clients might optimize different losses according to $\tau(c)$, which will produce diverging gradients across tasks.  
\end{itemize}  
In addition, systems remain heterogeneous, stale updates are propagated by client availability and update frequency. Therefore, the optimization problem must include partial participation and nonuniform local progress. Let $\ell_{\tau(c)}(\theta_e,\theta_{\tau(c)};\xi)$ be the loss cost for the specific sample $\xi\in\mathcal{D}_c^t$. The universal goal at stage $t$ is  
\begin{equation}  
\begin{aligned}  
\min_{\theta_e,\{\theta_k\}} \,F_t(\Theta)  
&= \sum_{c=1}^{C} p_c^t \, \mathbb{E}_{\xi\sim\mathcal{D}_c^t}  
\big[\ell_{\tau(c)}(\theta_e,\theta_{\tau(c)};\xi)\big] \\  
&\quad + \lambda \, \Omega(\{\theta_k\}),  
\end{aligned}  
\end{equation}  
where $p_c^t$ denote the weight of the aggregated model (i.e., the weight per unit local sample), $\Omega$ is a regularizer to enhance the cross-task learning stability and $\lambda\ge0$ moderates the sharing-specialization tradeoff. After local updates, the server also aggregates common encoder parameters through weighted averaging,  
\begin{equation}  
\theta_e^{t+1} = \sum_{c\in\mathcal{C}_t} \alpha_c^t \, \theta_{e,c}^{t+1}.  
\end{equation}  
and updates custom task-specific parameters, following task-wise aggregation,  
\begin{equation}  
\theta_k^{t+1} = \sum_{c \in\mathcal{C}_t: \, \tau(c)=k} \beta_c^t \, \theta_{k,c}^{t+1}.  
\end{equation}  
In $\mathcal{C}_t$ the client subset participates at round $t$ and $\alpha_c^t,\beta_c^t$ can be viewed as normalized aggregation coefficients \cite{mcmahan2017fedavg,smith2017mocha,zhou2025mfed}. By working towards privacy and heterogeneity requirements, RT-FedMTL is captured as the architecture in real-time, jointly maximizing the optimal use of shared representations and task-specific parameters (3, 4).