% \section{INTRODUCTION}
% \label{sec:introduction}

% With the proliferation of NLU systems on mobile devices, enterprise platforms, and edge services, the demand for privacy-preserving collaborative learning has grown significantly \cite{li2020challenges}. Centralized collection of NLU data from sources such as personal messages, medical text, or behavioral logs is often restricted by strict privacy regulations and user concerns. Consequently, FL has emerged as a key solution by keeping raw data local and sharing only model updates \cite{mcmahan2017fedavg}. Modern NLU applications, ranging from sentiment analysis for customer feedback to real-time intent classification for conversational agents and semantic similarity for information retrieval, increasingly require the ability to learn from diverse, distributed data streams while supporting MTL simultaneously \cite{smith2017mocha,chen2024fedbone}. The trend makes centralized single-pipeline training less feasible and encourages FL training models, which can learn collaboratively without breaking the data locality.

% While FL is a promising technique for NLU, there are three major concerns, as highlighted by benchmarking frameworks like FedNLP \cite{lin2021fednlp}. First, client data usually is non-IID. Given that diverse users and organizations write language data with different vocabularies, label compositions and preferences, convergence is typically slowed, and global model quality is poor \cite{li2020fedprox,smith2017mocha}. Second, task heterogeneity is common in realistic deployments. A federation could be made with intent classification, sentiment analysis, question answering, or even generation-related goals all at once \cite{smith2017mocha,chen2024fedbone,zhou2025mfed}. Classical one-model FL methods are not suited to this mixed-task regime. Third, communication and scalability are essential. In general, the repeated syncing of multiple clients can lead to significant overhead, particularly when the size of the model is large and the connected clients unstable \cite{mcmahan2017fedavg,chen2024fedbone}. Recent literature on federated MTL (e.g., MOCHA \cite{smith2017mocha}, FedBone \cite{chen2024fedbone}, MIRA \cite{elbakary2025mira}, and encoder-decoder based M-Fed \cite{zhou2025mfed}) indicates that data heterogeneity and task heterogeneity need to be considered as first-class design considerations.

% SLMs are especially appealing in federated systems because they improve efficiency without losing language pretraining functionality. Key advantages include:
% \begin{itemize}
%     \item Communication efficiency: A smaller parameter set reduces the communication cost per round (both upload and download).
%     \item Reduced computation cost: SLMs use fewer resources of local memory and training time, allowing for broader client involvement.
%     \item Edge-device deployment: compact models are easier to deploy and update on resource-constrained devices.
% \end{itemize}
% This role translates into parameter-efficient adaptation techniques (e.g., adapter-style and low-rank updates) that preserve competitive performance while decreasing federated overhead.

% Current FL pipelines predominantly rely on batch-based synchronization, where clients execute multiple local steps and synchronize periodically \cite{mcmahan2017fedavg,li2020fedprox}. While robust for static datasets, this design is less suited for streaming and dynamic environments, where language patterns, subjects, and user intent evolve over time. To address these limitations, this paper proposes RT-FedMTL, a real-time FL-MTL framework specifically designed for low-latency synchronization and agile adaptation. By utilizing WebSocket protocols and optimizing for SLMs, RT-FedMTL reduces the feedback loop, enabling near-instantaneous global updates that are essential for responsive applications like chatbots, live moderation, and event-driven analytics.

% This work makes several key contributions to the field of federated NLU:
% \begin{itemize}
%     \item We introduce RT-FedMTL, a novel real-time federated framework that integrates MTL and SLMs to leverage cross-task knowledge transfer while minimizing on-device resource footprints.
%     \item We evaluate five SLM architectures (DistilBERT \cite{sanh2019distilbert}, BERT-Medium \cite{devlin2018bert}, BERT-Mini \cite{devlin2018bert}, MiniLM \cite{wang2020minilm}, and TinyBERT \cite{jiao2019tinybert}) across three benchmark NLU tasks (sentiment analysis, paraphrase detection, and semantic similarity) to identify the optimal balance between performance and resource consumption.
%     \item We analyze the impact of non-IID and task heterogeneity on convergence speed and task accuracy in a real-time environment.
%     \item We demonstrate that RT-FedMTL achieves a significant reduction in per-client GPU memory usage (up to 74-93\%) compared to centralized MTL baselines while maintaining high performance across multiple benchmarks from the GLUE suite.
% \end{itemize}

% The rest of the paper is structured as follows. Section~\ref{sec:related_work} reviews the relevant work in FL and MTL. Section~\ref{sec:problem_definition} defines the formal problem of RT-FedMTL. In Section~\ref{sec:proposed_method} we detail the system architecture and optimization strategies. The experimental setup and results analysis are described in Section~\ref{sec:experiments}. Finally, Section~\ref{sec:conclusion} wraps up the paper and presents the next steps.

\section{Introduction} \label{sec:introduction} \IEEEPARstart{T}{he} prevalence of NLU systems on mobile devices, enterprise platforms, and edge services led to the development of an extensive and popular trend towards privacy-preserving collaborative learning \cite{li2020challenges}. Centralized aggregation of NLU data from sources like personal messages, medical text, or behavioral logs is typically limited by serious privacy laws and user concerns. Thus, FL has proven to be a promising approach by saving raw data locally and providing only updates to the model \cite{mcmahan2017fedavg}. Current NLU solutions like sentiment analysis for customer feedback, real-time intent classification for conversational agents, or semantic similarity for information retrieval are demanding to learn from various, distributed data streams while supporting MTL at the same time \cite{smith2017mocha,chen2024fedbone}. This tendency leads to the reduction of single-pipeline training in a centralized fashion becoming increasingly impractical and encourages FL training models, which can learn in a cooperative manner while preserving the data locality. Although FL is an innovative approach for NLU, three important limitations have already been identified based on benchmarking frameworks such as FedNLP \cite{lin2021fednlp}. For the starting point, client-generated data is generally non-IID. Because language data from multiple users and organizations come with different vocabularies, label compositions and preferences, convergence is often hindered, while global model quality is poor \cite{li2020fedprox,smith2017mocha}. Task heterogeneity is common in realistic deployments, second. We could achieve federation all at once such as intent classification, sentiment analysis, question answering, or even generation goals \cite{smith2017mocha,chen2024fedbone,zhou2025mfed}. Classical one-model FL methods fail in this mixed task regime. Third, you also need communication and scalability. Generally, for the larger size of models and unstable inter clients, the additional synchronization of multiple clients can cause a significant overhead \cite{mcmahan2017fedavg,chen2024fedbone}. Recent reports on federated MTL (e.g. MOCHA \cite{smith2017mocha}, FedBone \cite{chen2024fedbone}, MIRA \cite{elbakary2025mira}, encoder-decoder based M-Fed \cite{zhou2025mfed}) demonstrate that data heterogeneity and task heterogeneity should be addressed as first-class design concerns. SLMs are particularly attractive in federated systems since they increase efficiency without affecting language pretraining. Key advantages include:
\begin{itemize}
    \item Communication efficiency: By having fewer set of parameters, the communication cost is diminished in per-round (both upload & download).
    \item Cost savings in computation: SLMs consume less local memory and training time, enabling larger client involvement.
    \item Edge-device deployment: smaller models enable better deployment and upgrades on resource-limited devices.
\end{itemize}
This contribution translates into parameter-optimized adaptation processes (e.g., adapter style updates and low-rank updates) which promote competitive performance while reducing federated overhead. Current FL pipelines mainly depend on batch synchronization where clients execute multiple local steps and will execute the sync periodically \cite{mcmahan2017fedavg,li2020fedprox}. This design is effective for static datasets but less so for streaming and dynamic environments where language patterns as well as subjects and user intent change with frequency. This paper proposes RT-FedMTL, a modern real-time FL-MTL framework, to address these limitations and for low-latency synchronization and agile adaptation. RT-FedMTL utilizes WebSocket protocols and SLMs tuning to reduce the feedback loop and provide near-instantaneous global updates which is vital for the more responsive applications in application including chatbots, live moderation, and event-driven analytics. Here we contribute with several important insights to federated NLU.
\begin{itemize}
    \item We propose RT-FedMTL, a new real-time federated framework of MTL and SLMs to harness cross-task knowledge transfer without causing a large amount of resources to be consumed on devices.
    \item We assess five SLM architectures (DistilBERT \cite{sanh2019distilbert}, BERT-Medium \cite{devlin2018bert}, BERT-Mini \cite{devlin2018bert}, MiniLM \cite{wang2020minilm}, and TinyBERT \cite{jiao2019tinybert}) with respect to three standard NLU tasks (sentiment analysis, paraphrase detection, and semantic similarity), to get the best trade-off between performance and resource usage.
    \item We investigate the effects of non-IID and task heterogeneity on convergence speed and task accuracy in real-time domain.
    \item We show RT-FedMTL can drastically minimize per-client GPU memory requirements (up to 74-93\%) versus centralized MTL baselines, while keeping performance for multiple benchmarks (GLUE suite).
\end{itemize}
The remainder of the paper is organised as follows. The relevant work in FL and MTL is reviewed in section~\ref{sec:related_work}. Section~\ref{sec:problem_definition} defines the formal problem of RT-FedMTL. Section~\ref{sec:proposed_method} presents the system architecture and optimization techniques. Section~\ref{sec:experiments} describes the experimental setup and results analysis. Also, Section~\ref{sec:conclusion} closes the manuscript and describes the next steps.